
# **Flint Water Crisis**


## Summary
#####This is a very small data set where I explore the relationship between chlorine, copper, and lead concentration. There is a clear relationship between copper and lead concentration. However, I will need to compliment this data to find more significant relationships between other compounds and elements.

## EDA
```{r, include=F}
library(dplyr)
library(mclust)
library(ROCR)
library("plyr")
library(dplyr)
library(ggplot2)
library('reshape2')
library('gbm')
library('ROCR')
library("e1071")
library("class")
library("ranger")
library("TDboost")
library("caret")
library("glmnet")
library("ROCR")
```
```{r}
#read in dataset
cilc<-read.csv('cilc.csv')
head(cilc)

```
```{r, include=F}
model_gbm1 <- train(
  Lead..uM.~.,
  tuneLength = 5,
  data = cilc, method = "gbm",
  trControl = trainControl(method = "cv", number = 5, verboseIter = TRUE)
)
```
```{r, eval=F}
model_gbm1 <- train(
  Lead..uM.~.,
  tuneLength = 5,
  data = cilc, method = "gbm",
  trControl = trainControl(method = "cv", number = 5, verboseIter = TRUE)
)
```
```{r}
# I ran a gbm to take a look at the relative influence of the variables on lead
head(summary(model_gbm1))


cilc1<-cilc
# Lets change continueous variables to categorical
# Use quartile ranges to make the categories
# category 1 being the first quartile etc
# Lead categories will be based off of the microparts per billion that are considered safe by the EPA
cilc1$Lead..uM.<-cilc$Lead..uM.*100
cilc1$Chlorine..mM.<-cilc$Chlorine..mM.*100
cilc1$pH<-cilc$pH*100

cilc1$Lead..uM.<-as.integer(cilc$Lead..uM.)
cilc1$Copper..uM.<-as.integer(cilc1$Copper..uM.)
cilc1$Chlorine..mM.<-as.integer(cilc1$Chlorine..mM.)
cilc1$pH<-as.integer(cilc1$pH)

plot(density(cilc1$Chlorine..mM.))
plot(density(cilc1$Copper..uM.))
plot(density(cilc1$Lead..uM.))
plot(density(cilc1$pH))

summary(cilc1)

#pb above 14ppbs is bad
#ifelse&mutate to change the variable from numeric to categorical
cilc3<-cilc1
cilc3<-mutate(cilc3, lead = ifelse(Lead..uM. %in% 00:13, "1",
                                     ifelse(Lead..uM. %in% 14:20, "2", "3")))

cilc3<-mutate(cilc3, lead_dich = ifelse(Lead..uM. %in% 00:13, "safe", "unsafe"))

#copper above 14ppbs is bad
#ifelse&mutate to change the variable from numeric to categorical
cilc3<-mutate(cilc3, copper = ifelse(Copper..uM. %in% 00:75, "1",
                                     ifelse(Copper..uM. %in% 76:160, "2",
                                            ifelse(Copper..uM. %in% 161:240, "3", "4"))))

cilc3<-mutate(cilc3, chlorine = ifelse(Chlorine..mM. %in% 00:47, "1",
                                     ifelse(Chlorine..mM. %in% 48:66, "2",
                                            ifelse(Chlorine..mM. %in% 67:78, "3", "4"))))

cilc3<-mutate(cilc3, pH = ifelse(pH %in% 00:723, "1",
                                     ifelse(pH %in% 724:729, "2",
                                            ifelse(pH %in% 730:737, "3", "4"))))



cilc3$lead<-factor(cilc3$lead)
levels(cilc3$lead)
cilc3$copper<-factor(cilc3$copper)
cilc3$lead_dich<-factor(cilc3$lead_dich)
cilc3$chlorine<-factor(cilc3$chlorine)
cilc3$pH<-factor(cilc3$pH)
summary(cilc3)




# For the most part higher copper means high lead.
ggplot(data=cilc3, aes(x=Lead..uM., fill=copper, colour=copper))+ geom_density()+scale_x_log10()+facet_wrap(~copper)+geom_rug()

# There is a group along the y axis where lead does not increase as copper increases
# I think this is because these houses have the lead pipes with copper fittings but the Chloride and pH levels are non corrosive

# In this graph we see this group along the y-axis is 
ggplot(data=cilc3, aes(Lead..uM., Copper..uM., col=copper))+ geom_point()+geom_jitter(width=.5,height=.5, alpha=.2)+xlim(0,600)+ylim(0,300)
# In this graph we see the high copper and high lead levels are purple which is the highest chlorine group
ggplot(data=cilc3, aes(Lead..uM., Copper..uM., col=chlorine))+ geom_point()+geom_jitter(width=.5,height=.5, alpha=.2)+xlim(0,600)+ylim(0,300)

# pH
ggplot(data=cilc3, aes(x=Lead..uM., fill=pH, colour=pH))+ geom_density()+scale_x_log10()+facet_wrap(~pH)+geom_rug()







library(cluster) 
library(fpc)
wssplot <- function(data, nc=8, seed=1234){
               wss <- (nrow(data)-1)*sum(apply(data,2,var))
               for (i in 2:nc){
                    set.seed(seed)
                    wss[i] <- sum(kmeans(data, centers=i)$withinss)}
                plot(1:nc, wss, type="b", xlab="Number of Clusters",
                     ylab="Within groups sum of squares")
}

cilck <- dplyr::select(cilc3, copper, lead, pH, chlorine)

wssplot(cilck)
# 2 or 3 clusters
fit.km3<-kmeans(cilck, 3, nstart = 25)
#seeing the clusters compared to different categorical variables
table(cilc3$lead, fit.km3$cluster)
table(cilc3$copper, fit.km3$cluster)
table(cilc3$pH, fit.km3$cluster)
table(cilc3$chlorine, fit.km3$cluster)

clusplot(cilck, fit.km3$cluster)




cilck <- dplyr::select(cilc3, copper, lead, pH)

wssplot(cilck)
# 2 or 3 clusters
fit.km3<-kmeans(cilck, 3, nstart = 25)
#seeing the clusters compared to different categorical variables
table(cilc3$lead, fit.km3$cluster)
table(cilc3$copper, fit.km3$cluster)
table(cilc3$pH, fit.km3$cluster)
table(cilc3$chlorine, fit.km3$cluster)

clusplot(cilck, fit.km3$cluster)




cilck <- dplyr::select(cilc3, copper, lead, chlorine)

wssplot(cilck)
# 2 or 3 clusters
fit.km3<-kmeans(cilck, 3, nstart = 25)
#seeing the clusters compared to different categorical variables
table(cilc3$lead, fit.km3$cluster)
table(cilc3$copper, fit.km3$cluster)
table(cilc3$chlorine, fit.km3$cluster)

clusplot(cilck, fit.km3$cluster)
```


## GLM
```{r}
cilc3$lead_yn<-"Yes"
cilc3$lead_yn[cilc3$lead_dich=="safe"]<-"No"
cilc3$lead_yn<-factor(cilc3$lead_yn)

train.perc<-.8
dat.size<-nrow(cilc3)
train.amount<-as.integer(dat.size*train.perc)
train.row.nums<-sample(dat.size, train.amount, replace = F)
train.set<-subset(cilc3[train.row.nums,])
test.rows<-setdiff(1:dat.size, train.row.nums)
test.set<-subset(cilc3[test.rows,])

fit<-glm(lead_yn~copper*chlorine, cilc3, family=binomial(logit))

class.threshold<-.5
p<-predict(fit, test.set, type="response")
true.labels<-test.set[,12]
pred.labels<-rep("No",length(true.labels))
pred.labels[p>class.threshold]="Yes"


num.incorrect.labels<-sum(true.labels!=pred.labels)
misc.rate<-num.incorrect.labels/length(true.labels)
misc.rate

pred<-prediction(p,true.labels)
perf<-performance(pred,'tpr','fpr')

plot(perf,main="ROC")
abline(0,1,lty=3)

perf<- performance(pred,'auc',fpr.stop=1)
auc<-slot(perf,"y.values")[[1]]
auc
```

## GLM with bagging
```{r}
library(foreach)
train<-train.set
test<-test.set
set.seed(10) 
f<-lead_yn~copper*chlorine
length_divisor<-7  
iterations<-15 
predictions<-foreach(m=1:iterations,.combine=cbind) %do% {  
training_positions <- sample(nrow(train), size=floor((nrow(train)/length_divisor)))  
train_pos<-1:nrow(train) %in% training_positions  
lm_fit<-glm(f, data=cilc3, family=binomial(logit))
predict(lm_fit,newdata=test,type="response")  
}  
p<-rowMeans(predictions)

class.threshold<-.5
true.labels<-test.set[,12]
pred.labels<-rep("No",length(true.labels))
pred.labels[p>class.threshold]="Yes"


num.incorrect.labels<-sum(true.labels!=pred.labels)
misc.rate<-num.incorrect.labels/length(true.labels)
misc.rate


pred<-prediction(p,true.labels)
perf<-performance(pred,'tpr','fpr')

plot(perf,main="ROC")
abline(0,1,lty=3)

perf<- performance(pred,'auc',fpr.stop=1)
auc<-slot(perf,"y.values")[[1]]
auc
```


## SVM classification
```{r}
library('e1071')

col<-c("copper", "chlorine", "lead")


# Now lets narrow parameters to search very small epsilon values
tune_out<-tune(svm, lead~copper*chlorine,
  data=train.set,
  kernel="radial",
   ranges = list(epsilon = seq(0,.2,0.01), cost = 2^(2:9))
  )
print(tune_out)
plot(tune_out)

# We don't actually need to hand select the best parameters
# Finding best model
bestmodel=tune_out$best.model
summary(bestmodel)

# Predicting Species
ypredict<-predict(bestmodel, test.set, type="Class")

table(ypredict, test.set$lead)
```
