---
22title: "water_crisis_large_data"
author: "greg"
date: "February 26, 2017"
output: html_document
---

```{r, include=F}
library(dplyr)
library(mclust)
library(ROCR)
library("plyr")
library(dplyr)
library(ggplot2)
library('reshape2')
library('gbm')
library('ROCR')
library("e1071")
library("class")
library("ranger")
library("TDboost")
library("caret")
library("glmnet")
library("ROCR")
```
####In the first water_crisis project I looked at lead, copper, iron, ph, and chlorine. 
####This was interesting data but it was an unfortunately small set
####Only 82 in length
####Here I will do similar analysis on a larger data set 1200 in length.
####However this set does not have cholrine or pH levels so I cannot account for corrosiveness.
```{r}
cilc<-read.csv('cilcc.csv')
cilc1<-cilc
# Looking at iron and copper rq
cor(cilc$Copper, cilc$Iron)
cor(cilc$Lead, cilc$Iron)
cor(cilc$Lead, cilc$Copper)
# possibly a relationship between copper and iron


#change to categorical data
#lets multiply by 100 to make the ranges easier to work with also as.int to rid decimal
cilc1$Iron<-cilc$Iron*1000
cilc1$Iron<-log(cilc1$Iron)
cilc1$Iron<-cilc$Iron*100

cilc1$Lead<-as.integer(cilc1 $Lead)
cilc1$Copper<-as.integer(cilc1$Copper)
cilc1$Iron<-as.integer(cilc1$Iron)

summary(cilc1)

#pb above 14ppbs is bad
#ifelse&mutate to change the variable from numeric to categorical
cilc3<-cilc1
cilc3<-mutate(cilc3, lead3 = ifelse(Lead %in% 00:13, "1",
                                     ifelse(Lead %in% 14:20, "2", "3")))

cilc3<-mutate(cilc3, lead_dich = ifelse(Lead %in% 00:13, "safe", "unsafe"))

cilc3<-mutate(cilc3, copper = ifelse(Copper %in% 00:07, "1",
                                     ifelse(Copper %in% 08:23, "2",
                                            ifelse(Copper %in% 24:62, "3", "4"))))

cilc3<-mutate(cilc3, iron = ifelse(Iron %in% 00:08, "1",
                                     ifelse(Iron %in% 09:14, "2", "3")))

cilc3$lead3<-factor(cilc3$lead3)
cilc3$copper<-factor(cilc3$copper)
cilc3$iron<-factor(cilc3$iron)
```
##Basic GLM
```{r}
library(foreach)
cilc3$lead_yn<-"Yes"
cilc3$lead_yn[cilc3$lead_dich=="safe"]<-"No"
cilc3$lead_yn<-factor(cilc3$lead_yn)

train.perc<-.8
dat.size<-nrow(cilc3)
train.amount<-as.integer(dat.size*train.perc)
train.row.nums<-sample(dat.size, train.amount, replace = F)
train.set<-subset(cilc3[train.row.nums,])
test.rows<-setdiff(1:dat.size, train.row.nums)
test.set<-subset(cilc3[test.rows,])

train<-train.set
test<-test.set
set.seed(10) 
f<-lead_yn~Copper*Iron
length_divisor<-7  
iterations<-15 
predictions<-foreach(m=1:iterations,.combine=cbind) %do% {  
training_positions <- sample(nrow(train), size=floor((nrow(train)/length_divisor)))  
train_pos<-1:nrow(train) %in% training_positions  
lm_fit<-glm(f, data=cilc3, family=binomial(logit))
predict(lm_fit,newdata=test,type="response")  
}  
p<-rowMeans(predictions)


class.threshold<-.5
true.labels<-test.set[,11]
pred.labels<-rep("No",length(true.labels))
pred.labels[p>class.threshold]="Yes"


num.incorrect.labels<-sum(true.labels!=pred.labels)
misc.rate<-num.incorrect.labels/length(true.labels)
misc.rate

pred<-prediction(p,true.labels)
perf<-performance(pred,'tpr','fpr')

plot(perf,main="ROC")
abline(0,1,lty=3)

perf<- performance(pred,'auc',fpr.stop=1)
auc<-slot(perf,"y.values")[[1]]
auc
```
##SMOTE GLM
####My findings suggests lead concetration can be predicted by copper and lead
####The unexplained varience in lead concentration is probabaly explained by missing ph, chlorine, and temperature data
```{r}
table(cilc3$lead_yn)
# unbalanced data
# lets smote this and see if we can still make accurate predictions
library(DMwR)

cilc3$lead_dich<-factor(cilc3$lead_dic)
cilc4<-SMOTE(lead_dich~copper*iron, cilc3, k=5, perc.over = 200)
# check resultstrain
table(cilc4$lead_dich)

train.perc<-.8
dat.size<-nrow(cilc4)
train.amount<-as.integer(dat.size*train.perc)
train.row.nums<-sample(dat.size, train.amount, replace = F)
train.set<-subset(cilc4[train.row.nums,])
test.rows<-setdiff(1:dat.size, train.row.nums)
test.set<-subset(cilc4[test.rows,])

train<-train.set
test<-test.set
set.seed(10) 
f<-lead_dich~copper*iron
length_divisor<-7  
iterations<-15 
predictions<-foreach(m=1:iterations,.combine=cbind) %do% {  
training_positions <- sample(nrow(train), size=floor((nrow(train)/length_divisor)))  
train_pos<-1:nrow(train) %in% training_positions  
lm_fit<-glm(f, data=cilc4, family=binomial(logit))
predict(lm_fit,newdata=test,type="response")  
}  
p<-rowMeans(predictions)


class.threshold<-.5
true.labels<-test.set[,8]
pred.labels<-rep("safe",length(true.labels))
pred.labels[p>class.threshold]="unsafe"


num.incorrect.labels<-sum(true.labels!=pred.labels)
misc.rate<-num.incorrect.labels/length(true.labels)
misc.rate

pred<-prediction(p,true.labels)
perf<-performance(pred,'tpr','fpr')

plot(perf,main="ROC")
abline(0,1,lty=3)

perf<- performance(pred,'auc',fpr.stop=1)
auc<-slot(perf,"y.values")[[1]]
auc
```
SVM
```{r}
cilc<-read.csv('cilcc.csv')
cilc1<-cilc

cilc1$Lead<-as.integer(cilc1 $Lead)

summary(cilc1)

#pb above 14ppbs is bad
#ifelse&mutate to change the variable from numeric to categorical
cilc3<-cilc1
cilc3<-mutate(cilc3, lead3 = ifelse(Lead %in% 00:13, "1",
                                     ifelse(Lead %in% 14:20, "2", "3")))
sum(is.na(cilc3$lead3))
class(cilc3$lead3)
factor(cilc3$lead3)
cilc3$lead3<-factor(cilc3$lead3)
train.perc<-.8
dat.size<-nrow(cilc3)
train.amount<-as.integer(dat.size*train.perc)
train.row.nums<-sample(dat.size, train.amount, replace = F)
train.set<-subset(cilc3[train.row.nums,])
test.rows<-setdiff(1:dat.size, train.row.nums)
test.set<-subset(cilc3[test.rows,])

library('e1071')

col<-c("Copper", "Iron", "lead3")
# Now lets narrow parameters to search very small epsilon values
tune_out<-tune(svm, lead3~Copper*Iron,
  data=train.set,
  kernel="radial",
   ranges = list(epsilon = seq(0,.2,0.01), cost = 2^(2:9))
  )
print(tune_out)
plot(tune_out)

# We don't actually need to hand select the best parameters
# Finding best model
bestmodel=tune_out$best.model
summary(bestmodel)

# Predicting Species
ypredict<-predict(bestmodel, test.set, type="Class")

table(ypredict, test.set$lead3)
# This model is predicting most of the safe lead levels correctly
# However most of the data is safe lead levels so the accuracy paradox makes these results not as significant
# We need to smote the data and re run this
```
